---
title: "class15 : genome informatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Comment everything out so that it doesn't install every single time 
# source("http://bioconductor.org/biocLite.R")
# biocLite()

# For this class, you'll also need DESeq2:
# ("DESeq2)
```

## RNA-Seq

First step is to read our countData and colData (metadata) files that we will use with DESeq

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
View(counts)
View(metadata)
```

Lets examine the metadata file to find **control** and **treated** columns (cell-lines)

```{r}
metadata$dex == "control"
```

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/4 
names(control.mean) <- counts$ensgene
```

```{r}
# what can you replace the divided by 4 with ? 
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/nrow(control)
names(control.mean) <- counts$ensgene
```

Do the same thing for the treated columns 
```{r}
treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds, ]
treated
treated.mean <- rowSums(counts[, treated$id]/nrow(treated))
names(treated.mean) <- counts$ensgene
```

we will combine our meancount data for bookeeping purposes
```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
```

Let's plot our data
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```
changing the plot to make it with log transformation 
```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log="xy")
```

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the heaed() or the View()function for example

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

### filter genes with zero expression
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
head(sort(to.rm))
mycounts <- meancounts[-to.rm,]
head(mycounts)
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```


```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

merging mycounts row.names x with anno ensgene as y
```{r}
mycounts.anno <- merge(mycounts, anno, by.x="row.names", by.y="ensgene")
View(mycounts.anno)
```

download these packages but comment out so it doesn't re download every time you load the Rmd
```{r}
# biocLite("AnnotationDbi")
# biocLite("org.Hs.eg.db")
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

```

```{r}
head(mycounts)
```

```{r}
mycounts$entrezid <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
head(mycounts[up.ind,])
```

```{r}
ysource("http://bioconductor.org/biocLite.R")
```
```{r}
biocLite("DESeq2")
```
## Figure out what this isn't working
```{r}
library(DESeq2)

```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```
```{r}
sizeFactors(dds)
dispersions(dds)
dds <- DESeq(dds)
results(dds)
```

```{r}
res <- results(dds)
res
summary(res)
```

```{r}
resOrdered <- res[order(res$pvalue),]
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```

```{r}
resOrdered <- res[order(res$pvalue),]
res01 <- results(dds, alpha=0.01)
summary(res01)
```

```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")

```

```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]

rownames(resSig01[i,])

plotCounts(dds, gene="ENSG00000103196", intgroup="dex")

d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)

boxplot(count ~ dex , data=d)

```

## Volcano plots 
Let’s make another commonly produced visualization from this data
```{r}
plot(res01$log2FoldChange, -log(res01$padj))
```

Make a colored version of the plot using a color vector
```{r}
mycols <- rep("black", nrow(res01))
mycols[abs(res01$log2FoldChange) > 2] <- "red"

inds <- (res01$padj < 0.01) & (abs(res01$log2FoldChange) > 2)
mycols[inds] <- "blue"

plot(res01$log2FoldChange, -log(res01$padj), col=mycols)
```

