---
title: "Structural Bioinformatics Class11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Read CSV file from Section 1 
```{r}
pdbstats <- read.csv("Data Export Summary.csv", row.names = 1)

```

Lets look at the table
```{r}
library(knitr)
kable(pdbstats)
```

Q.1A Determine the percentage of structure solved by X-Ray and Electron Microscopy. 

```{r}
# total number of entries
ntotal <- sum(pdbstats$Total)

# Percentage of each method 
pereach <- pdbstats$Total / ntotal

# rounding the values
percents <- round(pereach*100, digits = 2)
# number of X ray structure [1] and number of EM strcuture [5] 
percents[1]
percents[3]
```
A1.A There are `r percents[1]` %X-Ray structures and `r percents[3]` %EM structures in the PDB database as of `r Sys.Date()`


```{r}
# Adding a percent column to the table 
nstats <- pdbstats
nstats$percents <- percents
kable(nstats)
```

Q.1B What proportion of structures are proteins? 

```{r}
protot <- sum(nstats$Proteins)
proppro <- protot/ntotal
proppror <- round(proppro*100, digits = 2)
```

A.1B The percent of enteries that are proteins is `r proppror` %

Q.2 How many HIV-1 proteases are in the PDB database

A.2 There are 1130 HIV-1 proteases currently in the PDB database. 

Section 3

open bio3D
```{r}
library(bio3d)

```

```{r}
pdb <-read.pdb("1hsg.pdb")
pdb
```

Note that the attributes (+ attr:) of this object are listed on the last couple of lines. To find the attributes = attributes(pdb)
```{r}
attributes(pdb)
```
```{r}
head(pdb$atom)

# Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]

# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]

# Which allows us to do the following 
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")

# Print a summary of the coordinate data in $xyz
pdb$xyz

# Examine the row and column dimensions
dim(pdb$xyz)

# Print coordinates for the first two atom
pdb$xyz[ 1, atom2xyz(1:2) ]

# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
ca.inds
```

## Aligning multiple structures 
```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A") 
files <- get.pdb(ids, split = TRUE)
```

```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)
# print pdbs
pdbs
```

```{r}
# find invariant core
core <- core.find(pdbs)

# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)

# Perform PCA
pc.xray <- pca(pdbs)

# plot the PCA
plot(pc.xray)
```

