---
title: "Class 8"
author: "Neha Patel"
date: "October 25, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering 

 Our first example with **kmeans()** function.
 
```{r}
# Example plot to see how Rmarkdown works
plot(1:10, typ="l")
```
 
 Back to kmeans...

```{r}
# Generate some example data for clustering 
tmp <- c(rnorm(30, -3), rnorm(30, 3))
x <- cbind(x=tmp, y=rev(tmp))


k <- kmeans(x, centers = 2, nstart = 20)
palette(c("blue", "green"))
plot(x, col = k$cluster)
points(k$centers, col="black", pch=20, cex=3)
```

```{r}
# Generate some example data for clustering 
tmp <- c(rnorm(30, -3), rnorm(30, 3))
x <- cbind(x=tmp, y=rev(tmp))


k3 <- kmeans(x, centers = 3, nstart = 20)
palette(c("blue", "green"))
plot(x, col = k3$cluster)
points(k3$centers, col="black", pch=20, cex=3)

k3$tot.withinss
k$tot.withinss
```
 
# Inspect/print the results

# Q. How many points are in each cluster?
      30

# Q. What ‘component’ of your result object details
#      - cluster size? 30
#      - cluster assignment/membership? "clustering vector"
#      - cluster center? (-3.015920, 2.671446)

# Plot x colored by the kmeans cluster assignment and
#      add cluster centers as blue points

# Q. Repeat for k=3, which has the lower tot.withinss?






# Hierarchical clustering in R 

Let's try out the **hclust()** function for Hierarchical clustering in R. This function needs a distance matrix as input! 

```{r}
d <- dist(x)
hc<- hclust(d)
plot(hc)
# Draw a line on the dendrogram
abline(h=8, col = "red")
# cut the tree to yield cluster membership vector
cutree(hc, h=8) 
```


### A more 'real' example of data clustering 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
palette(c("blue", "green", "red"))
plot(x, col=col, pch=20)
# Q. Use the dist(), hclust(), plot() and cutree()

#      functions to return 2 and 3 clusters
d2 <- dist(x)
hc2 <- hclust(d2)
plot(hc2)
# Use cutree() functions to return 2 and 3 clusters
cluster3 <- cutree(hc2, k=3)
cluster2 <- cutree(hc2, k=2)
table(cluster3)
table(cluster2)
# Q. How does this compare to your known 'col' groups?
table(col)
table(cluster3, col)
```
# Principal component analysis 

```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
##   and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
                       paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
wt.values <- rpois(5, lambda=sample(x=10:1000, size=1)) 
ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))
mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

```{r}
head(t(mydata))
```


Now lets try to find "structure" in this data with **promt()**

```{r}
pca <- prcomp(t(mydata), scale = TRUE)
pca
```
Make a PC plot of PC1 vs PC2. For this I use the '$x' component of our output (i.e. pca$x).
```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per

barplot(pca.var.per, main="Scree Plot",
        xlab="Principal Component", ylab="Percent Variation")
# A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```



